---
title: "Mythbusters: back of a hand"
output: html_document
---
## Part 1
In episode 5 of season 13 of Mythbusters they tested the expression "know something like the back of one's hand". 
12 subjects tried to select a photo of their hand from 10 similar pictures.

11 of 12 subjects selected their own hand. 

$H_0$: subjects select pictures at random ($p=0.1$).

$H_1$: subjects are able to select a picture of their own hand better than at random ($p>0.1$).

Exact binomial test to test $H_0$ against $H_1$
```{r, echo=FALSE}
binom.test(11, 12, p = 0.1, alternative = "greater")
```
What kind of confidence interval does it provide?

Likelihood ratio based 95% confidence interval (`prop.test`)
```{r, warning=FALSE}
prop.test(11, 12, conf.level=0.95)$conf.int
```

**Get 95% Wilson's confidence interval. (`scoreci` from "PropCIs" or `binconf` from "Hmisc")**
```{r}
library("PropCIs")
scoreci(11, 12, 0.95)
```

**Calculate p-values for Wald and score tests. Compare all the results. **
```{r}
num_successes = 11
num_experiments = 12
pi_hat = num_successes/num_experiments
pi_0 = 0.1
z_statistic = (pi_hat - pi_0)/sqrt(pi_hat*(1-pi_hat)/num_experiments)
p <- 1 - pnorm(z_statistic)



pi_hat_2 = 7/12
pi_hat_gen = mean(pi_hat_2, pi_hat)
pi_hat_gen
z_twosamples = (pi_hat_2 - pi_hat)/sqrt(pi_hat_gen*(1-pi_hat_gen)*(1/6))

z_twosamples

1-pnorm(z_twosamples)
```
 
## Part 2
The same experiment was run with the pictures of palms. 7 of 12 subjects selected a picture of their own palm. 

$H_0$: subjects select pictures at random ($p=0.1$).

$H_1$: subjects are able to select a picture of their own palm better than at random ($p>0.1$).

```{r}
binom.test(7, 12, p=0.1, alternative = "greater")
```

Could we claim that everyone knows a back of their hand better that a palm?
Let's compare the results of the experiments.
Those are the same subjects, but we have no information in one-to-one correspondence of their answers. 

** Calculate a p-value for independent two-sample z-test using one-sided alternative.**
```{r}
# your code here
```

**Calculate 95% Wilson's confidence interval (use `diffscoreci` from "PropCIs"), and invert it with "invertCI.R".**
```{r}
# your code here
```

**Can we still use paired sample test? How many different configurations of one-to-one correspondence are possible? Calculate p-values and confidence intervals using them.**
```{r}
# your code here
```