---
title: "Lifespan and diet for rats"
output: html_document
---
195 lab rats were part of the experiment.  
106 of them were randomly assigned to a restricted diet, 89 were able to eat ad libitum. 
Dataset: measured lifespan for each animal from both groups.
Does the diet influence the lifespan?

```{r echo=FALSE, message=FALSE}
install.packages('dplyr')
install.packages('ggplot2')
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
```

195 lab rats were part of the experiment.  
106 of them were randomly assigned to a restricted diet, 89 were able to eat ad libitum. 
Dataset: measured lifespan for each animal from both groups.
Does the diet influence the lifespan?

```{r}
rats <- read.table("rats.txt", sep="\t", header=T)
```

```{r}
rats %>%
  ggplot(aes(x=lifespan)) +
  geom_histogram(aes(fill=diet, color=diet), alpha=0.5, position = 'identity', bins=30) +
  xlab("Lifespan in days") 
```

**Make a boxplot of the data. **
```{r}
rats%>%
  ggplot(aes(x=lifespan, y=diet))+
  geom_boxplot(aes(fill=diet, color=diet)) + 
  xlab('boxplot')
```

Are the samples normal? Let's check the q-q plots.
```{r, fig.height=5, fig.width=5}
rats %>%
  ggplot(aes(sample = lifespan, colour = diet)) +
  stat_qq() +
  stat_qq_line() 
```
In the first sample the tails are quite heavy, and in the second a couple of smallest observations deviate far from the line.

A biologist suggested we exclude rats that lived less than 400 days from the analysis because they are likely died from some illnesses not related to the diet. 

** Exclude the observations, check the q-q plots again **
```{r}
rats%>%
  filter(lifespan >= 400) + 
  ggplot(aes(sample = lifespan, colour = diet)) +
  stat_qq() +
  stat_qq_line() 

```
Let's test the null hypothesis $H_0$: average lifespan of rats does not change with the dietary restrictions against two-sided alternative.

** Test the hypothesis with the appropriate test.**
```{r}
rats_filtered <- rats%>%
  filter(lifespan>400)
library('PropCIs')
diffscoreci()
rats_filtered%>%
  filter(diet=='restricted')
t.test(rats_filtered$lifespan[rats_filtered$diet =='restricted'],
       rats_filtered$lifespan[rats_filtered$diet =='ad libitum'])
```

Would the results be very different had we not filtered the dataset? 
** Use the same test on unfiltered data.**
```{r}
t.test(rats$lifespan[rats$diet =='restricted'],
       rats$lifespan[rats$diet =='ad libitum'])
```

**************
Weindruch R., Walford R.L. **The Retardation of Aging and Disease by Dietary Restriction.**  Springfield: Charles C. Thomas, 1988.
