---
title: "Speed of light"
output: html_document
---
```{r echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
```

The following dataset consists of the results of a classic Michelson experiment (1880) - speed of light was measured with a rotating mirror. There are 100 observations:
```{r}
speed <- scan("speed.txt")

ggplot() + 
  aes(x=speed) + 
  geom_histogram(color='black', fill='red', bins=30) +
  xlab('Measured speed of light, km/h') 
```

Are they normally distributed?

Ecdf and the best fitting normal cdf in the same axis:
```{r}
data.frame(x=speed) %>%
  ggplot(aes(x=x)) +
  stat_ecdf(col='blue') +
  stat_function(fun = pnorm, args = list(mean=mean(speed), sd=sd(speed)), col='red')  +
  xlab('Measured speed of light, km/h') +
  ylab('Empirical and fitted normal\ndistribution functions')
```

**Plot nonparametric sample density estimate and the best fitting normal density in the same axes (use `ggplot`, `geom_density`, `stat_function`).**
```{r}
data.frame(x=speed) %>%
  ggplot(aes(x=x)) +
  geom_density(col='blue') +
  stat_function(fun = dnorm, args = list(mean=mean(speed), sd=sd(speed)), col='red')  +
  xlab('Measured speed of light, km/h') +
  ylab('Empirical and fitted normal\ndistribution functions')
  
?geom_density
```

**Do the measurements look normal on q-q plot? (try `stat_qq`, `stat_qq_line`)**
```{r, fig.height=7, fig.width=7}
data.frame(x=speed) %>%
  ggplot(aes(sample=speed)) +
  stat_qq() +
  stat_qq_line() 
# your code here
```

**Check whether the measurements are normal using statistical test**
```{r}
shapiro.test(data.frame(x=speed)$x)
#data is normal
```

**************
Michelson A.A. **Experimental determination of the velocity of light made at the U.S. Naval Academy, Annapolis.** Astronomical Papers, 1, 109-145, 1880.
