---
title: "Bullshit"
output: html_document
---

```{r echo=F}
addtrend <- function(x, y){
  y <- y[order(x)]
  x <- sort(x)
  smoothed <- loess.smooth(x, y)
  with(smoothed, lines(x, y, col = "blue", lwd=3))
}

PlotBulshitVs <- function(y.name) {
  x <- d$MeanBullshit
  y <- d[, y.name]
  rm <- which(is.na(x) | is.na(y))
  if (length(rm) > 0) {
    x <- x[-rm]
    y <- y[-rm]
  }
  p.val <- cor.test(x,y,method="spearman")$p.value
  par(mar=c(4.3,4.3,2.1,0.1) + .1, mfrow=c(1,1))
  plot(x, jitter(y, amount=0.1), pch=19, col=rgb(255,0,0,100,maxColorValue = 255),
       main=paste0('ρ=', round(cor.test(x,y,method="spearman")$estimate, 3), ', p=', round(cor.test(x,y,method="spearman")$p.value, 4)),
       xlab="Bullshit Receptivity", ylab=y.name)
  m <- lm(y~x)
  lines(c(0,6), m$coefficients[1] + m$coefficients[2] * c(0,6), lwd=3)
  addtrend(x,y)
}
```

Original study: http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0153419

"Bullshit is used as a technical term, defined as communicative expression that lacks plausibility and truth." Example:

* Hidden meaning transforms unparalleled abstract beauty.
* Imagination is inside exponential space time events.

## Data
```{r echo=F}
d <- read.csv("bullshit.csv")
```

Sample: 196 US citizens (43.4% women, 36.4 years old on average) recruited on Amazon Mechanical Turk.

Assignment:

1. Rate the profoundness of each statement on a 5-point Likert-scale ranging from 1 - not at all profound, through 2 - somewhat profound, 3 - fairly profound, 4 - definitely profound, to 5 - very profound.

2. Rate the favorableness of three potential Democratic candidates and three Republican candidates for US president on a 5-point Likert scale ranging from 1 - very unfavorable, through 2 - somewhat unfavorable, 3 - neither unfavorable nor favorable, 4 - somewhat favorable, to 5 - very favorable. 

3. Estimate theirself political liberalism/conservatism using 7-point Likert scale (1 - liberal, 7 - conservative).

Some of the statements for the evaluation were bullshit, some - simple mundane statements (e.g., "A wet person does not fear the rain").

## Analysis

* For every subject the average bullshit receptiveness was calculated.

* Correlations of bullshit receptiveness with conservativeness and favourableness values were estimated with Spearman coefficient.

* To test if it's significantly different from 0, t-test was used.

## Results

There are siginificantly large positive correlations between receiptiveness to bullshit and:

* favorable views of Cruz, Rubio, Trump

* conservatism.

## Raw data
```{r fig.height=5, fig.width=5, warning=F, echo=F}
PlotBulshitVs("Conservatism")
PlotBulshitVs("FavorablenessCruz")
PlotBulshitVs("FavorablenessRubio")
PlotBulshitVs("FavorablenessTrump")
PlotBulshitVs("FavorablenessClinton")
PlotBulshitVs("FavorablenessSanders")
PlotBulshitVs("FavorablenessOMalley")
```

## Criticism

* Sample is not representative of the population
* Unbalancedness:

```{r echo=F}
hist(d$Conservatism, breaks=seq(0.5, 7.5, by=1), col='red', main='', xlab='Conservatism')
box()
```

* One of the features is not continuous

Overall, plots and coefficients tell different stories.

******************
Pfattheicher, S., & Schindler, S. (2016). Misperceiving bullshit as profound is associated with favorable views of Cruz, Rubio, Trump and conservatism. PLoS ONE, 11(4), 1–7. 