---
title: "Faith and education"
output: html_document
---
General Social Survey — an annual US citizens survey with about 5000 questions. From the survey of 2008 we selected the data on faith and education levels for 2000 respondents. Are these two variables independent?

```{r}
library(dplyr)
library(ggplot2)
install.packages("tidyr")
library(tidyr)
install.packages("ggmosaic")
library(ggmosaic)
data <- matrix(c(9,  8,27, 8, 47,236,
                 23,39,88,49,179,706,
                 28,48,89,19,104,293),
               nrow=3, byrow=TRUE, 
               dimnames = list(c("Less than high school", "High school/junior college", "Bachelor/graduate"), 
                               c("don't", "no way", "some power", "sometimes", "doubts", "know")))

data
```

```{r}
data_long <- data %>%
  as.data.frame() %>%
  mutate(Education = row.names(data)) %>%
  pivot_longer(!Education, names_to = 'Beliefs', values_to='n') %>%
  mutate(Beliefs = factor(Beliefs, levels = c("don't", "no way", "some power", "sometimes", "doubts", "know")))

data_long <- as.data.frame(lapply(data_long, rep, data_long$n)) %>% dplyr::select(-n)

ggplot(data_long) +
  geom_mosaic(aes(x = product(Education, Beliefs), fill = Education))+
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  xlab('') + ylab('') +
  guides(fill="none")
```


** Test the independence with chi squared test ** 
```{r}
chisq_test <-chisq.test(data)
chisq_test
chisq.test(data, correct=F)$stdres
```

** Try using Fisher's exact test**
```{r, error=TRUE}
fischer_test<- fisher.test(data, simulate.p.value=TRUE)
fischer_test

```

**Calculate Cramer's V and gamma coefficients (`assocstats` from `vcd` and `GKgamma` from `vcdExtra`) **  
```{r, message=FALSE, warning=FALSE}
install.packages('vcd')
install.packages('vcdExtra')
library(vcd)
library(vcdExtra)
assocstats(data)
GKgamma(data)
```

****************
2008 General Social Survey, National Opinion Research center (Agresti A. **Categorical Data Analysis**. — Hoboken: John Wiley & Sons, 2013, table 3.2)
