---
title: "Freshmen academic performance"
output: html_document
---

```{r message=F}
library(dplyr)
library(ggplot2)
library(GGally)
theme_set(theme_bw())
```

A researcher has collected data on three psychological variables and four academic variables (standardized test scores) for 600 college freshman. 

```{r fig.height=8, fig.width=8}
data <- read.table("mmreg.csv", sep=',', header=T)
psych <- data[, 1:3]
ggpairs(psych)
ggpairs(data)
```

** Calculate pairwise Pearson and Spearman correlation coefficients and test whether they are different from 0. Use `rcorr` from `Hmisc`, and try plotting correlation matrices with `corrplot`.**
```{r, warning=FALSE, message=FALSE}
install.packages('Hmisc')
library(Hmisc)
rcorr(as.matrix(data), type="pearson")
rcorr(as.matrix(data), type="spearman")

install.packages('corrplot')
library(corrplot)
corr_df <- data.frame(rcorr(as.matrix(data))[1])
corrplot(corr_df)
corrplot(cor(data))
```

**For a selected pair of features, calculate Kendall's correlation coefficient ant its significance using `cor` and `cor.test`:**
```{r}
```

**Implement permutation test for Spearman correlation coefficient between math and self concept**
```{r}
install.packages('jmuOutlier')
library(jmuOutlier)
perm.cor.test(data$math, data$self_concept, method='spearman')
```

**Compare the result to `spearman_test` from `coin`**
```{r}
install.packages('coin')
library(coin)
spearman_test(math ~ self_concept, data=data)
?spearman_test
perm.cor.test
```

** With `pcor` from `ppcor` calculate Pearson's partial correlations (excluding the influence of everything except the pair) and compare them to simple pairwise correlations.  **
```{r}
```


******************
R Data Analysis Examples: Canonical Correlation Analysis. UCLA: Statistical Consulting Group. 
https://stats.idre.ucla.edu/r/dae/canonical-correlation-analysis/ 